---
title: "lab5_part1"
author: "Danielle Sclafani"
date: "2/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tsibble)
library(feasts)
library(fable)
library(here)
library(janitor)

```
### Part A & B

```{r}
#reading in the data

energy <- read_csv(here("data", "energy.csv"))

# converting the month column that is currently "March 2020" to a time series that R can read using tsibble

energy_ts <- energy %>% 
  mutate(date = tsibble::yearmonth(month)) %>% 
  as_tsibble(key = NULL, index = date)

```

### Part C: Exploring the time series data

```{r}
ggplot(data = energy_ts, aes(x = date, y= res_total)) +
  geom_line()+
  labs(y = "Residential Energy consumption \n (Trillion BTU)")
```


Looks like there are some interesting things happening. We should ask:

Is there an overall trend?
Is there seasonality?
Any cyclicality evident?
Any other notable patterns, outliers, etc.?

The big ones to notice quickly here are:

Overall increasing trend overall, but stability (and possibly a slight decreasing trend) starting around 2005
Clear seasonality, with a dominant seasonal feature and also a secondary peak each year - that secondary peak has increased substantially
No notable cyclicality or outliers

### Seasonplot using feasts::gg_season()

```{r}
energy_ts %>% 
  gg_season(y= res_total) +
  theme_minimal()+
  labs(x = "month",
       y = "residential energy consumption (trillion BTU)")
```



This is really useful for us to explore both seasonal patterns, and how those seasonal patterns have changed over the years of this data (1973 - 2017). What are the major takeaways from this seasonplot?

The highest residential energy usage is around December / January / February
There is a secondary peak around July & August (that’s the repeated secondary peak we see in the original time series graph)
We can also see that the prevalence of that second peak has been increasing over the course of the time series: in 1973 (orange) there was hardly any summer peak. In more recent years (blue/magenta) that peak is much more prominent.

#### Subseries plot

```{r}
energy_ts %>% gg_subseries(res_total)
```

Our takeaway here is similar: there is clear seasonality (higher values in winter months), with an increasingly evident second peak in June/July/August. This reinforces our takeaways from the raw data and seasonplots.

#### Decomposition by STL - STL allows for seasonality to change over time. which we need to use here because we are seeing that it is changing over time with the second summer peak increaseing over time. classical decomposition does not allow for seasonality to vary over time

```{r}
# finding STL decomp
dcmp <- energy_ts %>% 
  model(STL(res_total ~ season()))

#visuaizing the components
components(dcmp) %>%  autoplot() +
  theme_minimal()

```

#### Autocorrelation Function - use the ACF to explore autocorrelation - here we expect seasonality to be clear from the ACF

```{r}
energy_ts %>% 
  ACF(res_total) %>% 
  autoplot
```

And yep, we see that observations separated by 12 months are the most highly correlated, reflecting strong seasonality we see in all of our other exploratory visualizations.


### Part D: Forecasting by Holt-Winters exponential smoothing
Note: here we use ETS, which technically uses different optimization than Holt-Winters exponential smoothing, but is otherwise the same (From Rob Hyndman: “The model is equivalent to the one you are fitting with HoltWinters(), although the parameter estimation in ETS() uses MLE.”)

To create the model below, we specify the model type (exponential smoothing, ETS), then tell it what type of seasonality it should assume using the season("") expression, where “N” = non-seasonal (try changing it to this to see how unimpressive the forecast becomes!), “A” = additive, “M” = multiplicative. Here, we’ll say seasonality is multiplicative due to the change in variance over time and also within the secondary summer peak:

```{r}
#creating the model
energy_fit <- energy_ts %>% 
  model(
    ets = ETS(res_total ~ season("M"))
  )

#forecasting using the model 10 years into the future

energy_forecast <- energy_fit %>% 
  forecast(h = "10 years")

#plot just the forecasted values with 80 & 95% Confidence intervals

energy_forecast %>% 
  autoplot

```

```{r}
# can also plot the forecasted data with the original data

energy_forecast %>% 
  autoplot(energy_ts)
```

#### Assessing Residuals
We can use broom::augment() to append our original tsibble with what the model predicts the energy usage would be based on the model. Let’s do a little exploring through visualization.

First, use broom::augment() to get the predicted values & residuals:
```{r}
# append the predicted values and residuals to original energy data
energy_predicted <- broom::augment(energy_fit)

```

```{r}
# plotting actual energy values and the predicted values

ggplot(data = energy_predicted) +
  geom_line(aes(x = date, y = res_total))+
  geom_line(aes(x = date, y = .fitted, color = "red"))

```

Alot of overlap = good, means predictions are pretty good

#### exploring the residuals - residuals should be uncorrelated, centered at 0 and ideally normaly distributed 

```{r}
# check residuals with a histogram
ggplot(data = energy_predicted, aes(x = .resid)) +
  geom_histogram()
```











