---
title: "lab5_part2"
author: "Danielle Sclafani"
date: "2/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tsibble)
library(feasts)
library(fable)
library(here)
library(janitor)
library(sf)

```

# Part 2: Spatial data wrangling, visualization, and a variogram

```{r}
ca_counties <- read_sf(here("data", "ca_counties","CA_Counties_TIGER2016.shp"))
```

```{r}
# wrangling

ca_subset <- ca_counties %>% 
  select(NAME, ALAND) %>% 
  rename(county_name = NAME, land_area = ALAND)

```

```{r}
# checking out to coordinate system 

ca_subset %>% st_crs()
```

```{r}
# plotting the spatial data
ggplot(data = ca_subset)+
  geom_sf(aes(fill = land_area), color = "white", size = 0.1) +
  theme_void()+scale_fill_gradientn(colors = c("cyan", "blue", "purple"))


```

## Part B. Invasive Red Sesbania records 

```{r}
#reading in data for the invasive species

sesbania <- read_sf(here("data", "red_sesbania", "ds80.shp"))

#checking the coordinates

sesbania %>% st_crs()


```


since the CRS is different for the map we made above and the sesbania data we want it to match
```{r}
#matching the coordinate systems

sesbania <- st_transform(sesbania, 3857)

sesbania %>% st_crs()
```

## combine the graphs

```{r}
ggplot() +
  geom_sf(data = ca_subset)+
  geom_sf(data = sesbania, size = 1, color = "red")
```

## joining data subsets
- we want to find the counts of sesbania in each county. first we have to join the subsets
```{r}
ca_sesbania <- ca_subset %>% 
  st_join(sesbania)
```

```{r}
#finding the counts of sesbania by county
sesbania_counts <- ca_sesbania %>% 
  count(county_name)
```


Then we can plot a chloropleth using the number of records for red sesbania as the fill color (instead of what we used previously, land area):


```{r}
ggplot(data = sesbania_counts) +
  geom_sf(aes(fill = n), color = "white", size = 0.1) +
  scale_fill_gradientn(colors = c("lightgrey", "orange", "red"))+
  theme_minimal()+
  labs(fill = "Number of S. punicea records")
```


So we see that we can still use our usual wrangling skills! Let’s do a bit more for fun, just to prove that our existing wrangling skills still work with spatial data - the spatial information just sticks to it!
Only plot the county with the greatest number of red sesbania records (Solano), and make a map of those locations (yeah there are many ways to do this):

```{r}
#subset of sesbania in Solano county

solano_sesbania <- sesbania %>% 
  filter(COUNTY == "Solano")

solano <- ca_subset %>% 
  filter(county_name =="Solano")

ggplot()+
  geom_sf(data = solano)+
  geom_sf(data = solano_sesbania)

```



### Part C - Making an interactive map


